---
title: "TAPIR_FINAL_GF_ONLY"
author: "Luisa Egert"
date: "2026-01-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# set pathways
```{r}
MainDIR_TP3="/Users/ra98rux/Desktop/RNA_Seq_MW/TAPIR_Anna_CountMTX"

WORKDIR_GF=file.path(MainDIR_TP3,"GF_ONLY_Analysis")
#dir.create(WORKDIR_GF)
ResultsDIR_TP3=file.path(WORKDIR_TP3, "Results")
#dir.create(ResultsDIR_TP3)
ResultsDIR_GF=file.path(WORKDIR_GF, "Results")
#dir.create(ResultsDIR_GF)
FigDIR_GF=file.path(WORKDIR_GF,"Plots")
#dir.create(FigDIR_GF)

CountsDIR_TP3=file.path(MainDIR_TP3,"CountFiles")

GeneScoreDIR=file.path(WORKDIR_TP3,"GeneScores")

```

# install libraries
```{r}
#BiocManager::install("rtools")
#install.packages("xfun")
#BiocManager::install("edgeR")
#BiocManager::install("biomaRt")
#BiocManager::install("statmod")
#BiocManager::install("ggrepel")
#BiocManager::install("umap")
#BiocManager::install("EnsDb.Hsapiens.v86")
#BiocManager::install("sva")
#BiocManager::install("rstatix")
#BiocManager::install("DESeq2")
#BiocManager::install("msigdbr")
#BiocManager::install("fgsea")
#BiocManager::install("tidyverse")
#BiocManager::install("tidyr")
#BiocManager::install("rtracklayer")
#BiocManager::install("gridExtra")
#BiocManager::install("grid")
# install biomart and genomicfeatures from bioconductor
#if (!requireNamespace("BiocManager", quietly = TRUE))
   # install.packages("BiocManager")

#BiocManager::install("biomaRt")
#BiocManager::install("GenomicFeatures")

library(edgeR)
library (ggplot2)
library(plyr)
library(biomaRt)
library(ggrepel)
library(dplyr)
library(pheatmap)
library(gplots)
library(cowplot)
library(sva)
library(rstatix)
library(DESeq2)
library(fgsea)
library(tidyverse)
library(msigdbr)
library(tidyr)
library(rtracklayer)
library(gridExtra)
library(grid)

library(biomaRt)
library(GenomicFeatures)


```

# make ens_mgi list with rtracklayer (with gtf file from anna d)
```{r}

# Import GTF file
gtf <- import(file.path(MainDIR_TP3, "gencode.vM35.annotation.gtf"))

# Extract unique gene info
genes <- unique(mcols(gtf[gtf$type == "gene", c("gene_id", "gene_name")]))
head(genes)

rownames(genes)<-genes$gene_id
#add ensvowers
genes$enswovers<-gsub("\\..*","",row.names(genes))
rownames(genes)<-genes$enswovers

# remove columns with ens with verisons
genes.1<-genes[,-3]

```

# read in Meta File
```{r}
Meta_TP3_wo3_GF<-read.table(file.path(ResultsDIR_TP3, "Meta_TP3_wo3_GF.txt"), sep="\t")
rownames(Meta_TP3_wo3_GF)<-Meta_TP3_wo3_GF$Sample_ID
```


# read in count files and make count matrix - GF 
```{r}
# create count table
countl_GF<-list.files(path = CountsDIR_TP3,pattern = ".txt", ignore.case = FALSE) #reading in the file names as a list

#a sorting vector according to RNAseq ID
sortvec_GF<-sapply(as.character(Meta_TP3_wo3_GF$Sample_ID), function(x) grep(x,countl_GF))
sortvec_GF
length(sortvec_GF)

# creat count df
counts_GF<-as.data.frame(matrix(),row.names = NULL)
for (i in sortvec_GF) {
rpgl_GF.t<-read.delim(file.path(CountsDIR_TP3,countl_GF[i]),sep = "\t",header = F, skip =  ,check.names = TRUE,row.names = NULL)
counts_GF<-cbind(counts_GF,rpgl_GF.t[1:nrow(rpgl_GF.t),3])
}

# remove empty column
counts_GF<-counts_GF[,(-1)]

#add ens numbers as row names
ens_MW.GF<-read.delim(file.path(CountsDIR_TP3,countl_GF[3]),sep = "\t", header = F, skip = ,check.names = TRUE,row.names = NULL)[1:nrow(rpgl_GF.t),1]
row.names(counts_GF)<-ens_MW.GF

#add ensvowers
counts_GF$enswovers<-gsub("\\..*","",row.names(counts_GF))
#check for duplicates
nrow(counts_GF[duplicated(counts_GF$enswovers,fromLast=TRUE),]) #0 duplicated ENS

#remove dups
rownames(counts_GF)<-counts_GF$enswovers 

colnames(counts_GF)<-Meta_TP3_wo3_GF$Sample_ID

counts_GF<-counts_GF[,-13]

#write.table(counts_GF,file=file.path(ResultsDIR_GF, "counts_GF.txt"),sep="\t")

counts_GF<-read.table(file = file.path(ResultsDIR_GF, "counts_GF.txt"), header = T,sep="\t")

```

# selec for genes in count df
```{r}
# select for genes in count df
genes2<-subset(genes.1, rownames(genes.1)%in%rownames(counts_GF))

#write.table(genes2, file=file.path(ResultsDIR_TP3, "genes2.txt"), sep="\t")

genes2<-read.table(file = file.path(ResultsDIR_TP3, "genes2.txt"), header = T,sep="\t")

```


## DEG with DESeq2 on raw counts 
```{r}
## design = ~ TAPIR + DS
dds.wo3_GF<-DESeqDataSetFromMatrix(
  countData = counts_GF, 
  colData = Meta_TP3_wo3_GF, 
  design = ~ DS + treadCond) # in deseq2: covariats first (batch) treatment last

dds.wo3_GF<-DESeq(dds.wo3_GF)
dds.wo3_GF.counts<-counts(dds.wo3_GF, normalized=TRUE)

# To get the main comparison of control vs gRNA, extract results
res_gRNAvsCtr.GF.wo3_ONLY<-results(dds.wo3_GF, contrast = c("treadCond", "GF_gRNA", "GF_CTR"))

## remove samples with N/A in padj column
res_gRNAvsCtr.GF.wo3_ONLY_woNA<-res_gRNAvsCtr.GF.wo3_ONLY[!is.na(res_gRNAvsCtr.GF.wo3_ONLY$padj), ]

## merge df wo NA filter with mgi
res_gRNAvsCtr.GF.wo3_ONLY_woNA<-as.data.frame(res_gRNAvsCtr.GF.wo3_ONLY_woNA)
res_gRNAvsCtr.GF.wo3_ONLY_woNA.mgi<-merge(res_gRNAvsCtr.GF.wo3_ONLY_woNA, genes2, by=0)
#write.table(res_gRNAvsCtr.GF.wo3_ONLY_woNA.mgi,file=file.path(ResultsDIR_GF, "res_gRNAvsCtr.GF.wo3_ONLY_woNA.mgi.txt"),sep="\t")


# remove genes with a base mean expression < 1.5
## check MA plots
plotMA(dds.wo3_GF, ylim=c(-15, 15), alpha = 0.1)

## filter for base mean
res_gRNAvsCtr.GF.wo3_ONLY_woNA_BM<-res_gRNAvsCtr.GF.wo3_ONLY_woNA[(res_gRNAvsCtr.GF.wo3_ONLY_woNA$baseMean >1.5),]

#write.table(res_gRNAvsCtr.GF.wo3_ONLY_woNA_BM,file=file.path(ResultsDIR_GF, "res_gRNAvsCtr.GF.wo3_ONLY__woNA_BM.txt"),sep="\t")


# order by different conditions
res_gRNAvsCtr.GF.ONLY_wo3_woNA.padj<-res_gRNAvsCtr.GF.wo3_ONLY_woNA[order(res_gRNAvsCtr.GF.wo3_ONLY_woNA$padj),] 
res_gRNAvsCtr.GF.ONLY_wo3_woNA.pval<-res_gRNAvsCtr.GF.wo3_ONLY_woNA[order(res_gRNAvsCtr.GF.wo3_ONLY_woNA$pvalue),] 
res_gRNAvsCtr.GF.ONLY_wo3_woNA.FLC<-res_gRNAvsCtr.GF.wo3_ONLY_woNA[order(res_gRNAvsCtr.GF.wo3_ONLY_woNA$log2FoldChange),] 

# make list with interesting genes
summary.df_GF<-data.frame()
summary.df_GF["GF_gRNA_vs_GF_CTR", "Up_Padj"]<-nrow(subset(as.data.frame(res_gRNAvsCtr.GF.ONLY_wo3_woNA.padj), padj<0.05 & log2FoldChange > 0.585))
summary.df_GF["GF_gRNA_vs_GF_CTR", "DN_Padj"]<-nrow(subset(as.data.frame(res_gRNAvsCtr.GF.ONLY_wo3_woNA.padj), padj<0.05 & log2FoldChange < (-0.585)))
summary.df_GF["GF_gRNA_vs_GF_CTR", "Up_PVAL"]<-nrow(subset(as.data.frame(res_gRNAvsCtr.GF.ONLY_wo3_woNA.padj), pvalue<0.05 & log2FoldChange > 0.585))
summary.df_GF["GF_gRNA_vs_GF_CTR", "DN_PVAL"]<-nrow(subset(as.data.frame(res_gRNAvsCtr.GF.ONLY_wo3_woNA.padj), pvalue<0.05 & log2FoldChange < (-0.585)))

summary.df_GF
#write.table(summary.df_GF,file=file.path(ResultsDIR_GF, "summary.df_GF.woNA.txt"),sep="\t")


# Use basemean filtered res tables and order
res_gRNAvsCtr.GF.wo3_ONLY_woNA_BM.padj<-res_gRNAvsCtr.GF.wo3_ONLY_woNA_BM[order(res_gRNAvsCtr.GF.wo3_ONLY_woNA_BM$padj),] 
res_gRNAvsCtr.GF.wo3_ONLY_woNA_BM.padj<-as.data.frame(res_gRNAvsCtr.GF.wo3_ONLY_woNA_BM.padj)

## merge df with BM filter with mgi
res_gRNAvsCtr.GF.wo3_ONLY_woNA_BM.padj.mgi<-merge(res_gRNAvsCtr.GF.wo3_ONLY_woNA_BM.padj, genes2, by=0)


```

## CBS on raw dds.counts_TP3.GF - batch correction for PCA
```{r}
batchvec_merge_GF<-Meta_TP3_wo3_GF$DS
raw_counts_GF <- counts(dds.wo3_GF, normalized = FALSE) # extract raw counts from DSEQ2 matrix
CBS_counts_TP3.wo3_GF<-ComBat_seq(as.matrix(raw_counts_GF), batch = batchvec_merge_GF) # batch correction

#write.table(CBS_counts_TP3.wo3_GF,file=file.path(ResultsDIR_GF, "CBS_counts_TP3.wo3_GF.txt"),sep="\t")

```

### dds.counts normalization on CBS treated raw counts - for PCA
```{r}
## design = ~ TAPIR + DS
dds.wo3.CBS_GF<-DESeqDataSetFromMatrix(
  countData = CBS_counts_TP3.wo3_GF, 
  colData = Meta_TP3_wo3_GF, 
  design = ~ DS + treadCond)

# isolate normalized counts
dds.wo3.CBS_GF<-DESeq(dds.wo3.CBS_GF)
dds.wo3.CBS_GF.counts<-counts(dds.wo3.CBS_GF, normalized=TRUE)

# write.table(dds.wo3.CBS_GF.counts,file=file.path(ResultsDIR_GF, "dds.wo3.CBS_GF.counts.txt"),sep="\t")

```

### PCA plot with outline
```{r}

group_TP3_wo.3.GF<-factor(Meta_TP3_wo3_GF$TAPIR, levels = c("gRNA", "control"))
outline_GF= c( "gRNA" = "darkred", "control" = "darkblue")

colourlables6<- c("gRNA", "control")
scale_colour6 = c( "gRNA" = "#FF6666", "control" = "#00CCFF")


pca_func.wo3.GF.s<-function(data, groupv = group_TP3_wo.3.GF, labels = "", celltype = group_TP3_wo.3.GF, colourlables = colourlables6, scales_col = scale_colour6, outline_col = outline_GF){
  prin_comp.d <- prcomp(data, scale. = FALSE)
  std_dev.d <- prin_comp.d$sdev 
  pr_var.d <- std_dev.d^2
  prop_varex.d <- round(100*pr_var.d/sum(pr_var.d), digits=1)
  embedding <- as.data.frame(prin_comp.d$rotation)
  embedding$Class <- as.factor(groupv)
  embedding$celltype <- as.factor(celltype)
  p1 <- ggplot(data=embedding) +
    aes(x=PC1, y=PC2)+
      geom_point( size = rel(16), stroke = 2, aes( fill = Class, colour = Class, shape = celltype)) +
      scale_colour_manual( values = outline_GF, name = "TAPIR GF", labels = colourlables) +
      scale_fill_manual( values = scales_col, name = "TAPIR GF",labels = colourlables) +
      scale_shape_manual( values = c("gRNA" = 22, "control" = 21),name = "TAPIR") +
      geom_text_repel(aes(label=labels), col="black", segment.colour="black", segment.size=0.4, min.segment.length=0.1, point.padding=.1,segment.alpha=0.7, alpha=1, size       = 4) +
      ggtitle("PCA: TAPIR 2")  +
      xlab(paste("PC2: ",prop_varex.d[2],"% of variance")) + ylab(paste("PC1: ",prop_varex.d[1],"% of variance")) +
theme_light() +
    theme(plot.tag=element_text(size = 12*2.0, face = "bold"),
          plot.title = element_text(size = rel(1.5), face = "plain", hjust = 0.5), 
          legend.text = element_text(colour="black", size = rel(1.2), face = "plain"),
          legend.title = element_text(colour="black", size = rel(1.2), face = "bold"),
          legend.box.just = "top",
          axis.text = element_text(colour = "black", size = rel(1), face = "plain"),
          axis.title = element_text(colour = "black",size = rel(1.5),face = "plain"),
          panel.grid = element_blank(),
          panel.border = element_rect(colour = "black",size=2),
          axis.ticks = element_line(colour="black"),
          aspect.ratio = 1.0, 

          legend.position = "right",legend.background = element_rect(fill = "transparent"),legend.box.background = element_rect(colour = "darkgrey")
        )
  plot(p1)
}

```

#### PCA top 500 diff genes for GF - FC - highest to lowest
```{r}
## GF
res_gRNAvsCtr.GF.wo3_ONLY_woNA_BM.o.FC.UP<-res_gRNAvsCtr.GF.wo3_ONLY_woNA_BM[order(res_gRNAvsCtr.GF.wo3_ONLY_woNA_BM$log2FoldChange, decreasing = TRUE),]
res_gRNAvsCtr.GF.wo3_ONLY_woNA_BM.o.FC.UP_top500<-res_gRNAvsCtr.GF.wo3_ONLY_woNA_BM.o.FC.UP[1:500,]

### subset counts with top500
top500FC.UP_CBS_GF_ONLY_wo3<-subset(dds.wo3.CBS_GF.counts, rownames(dds.wo3.CBS_GF.counts)%in%rownames(res_gRNAvsCtr.GF.wo3_ONLY_woNA_BM.o.FC.UP_top500))

## run PCA
pca_func.wo3.GF.s(top500FC.UP_CBS_GF_ONLY_wo3)

pdf(file=file.path(FigDIR_GF, "PCA_top500FC.UP_CBS_wo3.pointsize_outline_GF_only.pdf"), height = 10, width = 10)
pca_func.wo3.GF.s(top500FC.UP_CBS_GF_ONLY_wo3)
dev.off()

```

## 3 GSEA - with DESQ2 Data with human sig
```{r}

MsicList<- vector(mode = "list", length = 0)
signatures<-c("H","C2","C3","C5","C6","C7","C8")
for (sig in signatures){
  MsicList[[sig]]<-msigdbr(species = "Mus musculus", category = sig)
  #Transform into list format requires for indexing
  MsicList[[sig]]<- split(x = MsicList[[sig]]$gene_symbol, f = MsicList[[sig]]$gs_name)
}


#transform results into ranks format
## GF
res_gRNAvsCtr.GF.wo3_ONLY_woNA.mgi <- res_gRNAvsCtr.GF.wo3_ONLY_woNA.mgi %>%
  dplyr::select(gene_name, stat) %>%
  na.omit() %>% distinct() %>% group_by(gene_name) %>%
  summarize(stat=mean(stat))

ranksGF.only<-deframe(res_gRNAvsCtr.GF.wo3_ONLY_woNA.mgi)

# for C2 C5 and H combined 
set.seed(8501)
fgsea_C2C5H_GF.only <- fgsea(pathways = c(MsicList[["C2"]],MsicList[["C5"]],MsicList[["H"]]),stats = ranksGF.only,eps = 0,minSize = 6, maxSize = 200,)
fgsea_C2C5H_GF.only.o <- fgsea_C2C5H_GF.only[order(fgsea_C2C5H_GF.only$padj),]

## NES means normalized enrichment score (e.g. GF vs control, if NES more the 1 then it's enriched in GF)
fgsea_C2C5H_GF.only.o.UP <- subset(fgsea_C2C5H_GF.only, padj<0.01 & NES > 1)
fgsea_C2C5H_GF.only.o.DN <- subset(fgsea_C2C5H_GF.only, padj<0.01 & NES < (-1))

# vector with all signatures of list
combpathways <- c(MsicList[["C2"]],MsicList[["C5"]],MsicList[["H"]])

# significant Cell cycle signatures 
# for data frame created above and that have a signficant padj
Cellcyclepaths_Sig_GF.only <- combpathways[subset(fgsea_C2C5H_GF.only.o[grep("CELL_CYCLE",fgsea_C2C5H_GF.only.o$pathway),], padj<0.05)$pathway] # df that contains only sig cell cycle pathways
cellcyle_GFGSEAplot.only <- plotGseaTable(Cellcyclepaths_Sig_GF.only, ranksGF.only,
              fgsea_C2C5H_GF.only.o, gseaParam = 0.5)                  
#ggsave(cellcyle_GFGSEAplot.only,file=file.path(FigDIR_GF,"CellCycle_GFGSEAplot_maxsize200_Clean_ONLY.pdf"),height=10,width=15)


# significant ribosome signature
RIBOSOMEpaths_Sig_GF.only <- combpathways[subset(fgsea_C2C5H_GF.only.o[grep("RIBOSOME",fgsea_C2C5H_GF.only.o$pathway),], padj<0.05)$pathway]
RIBOSOME_GFGSEAplot.only <- plotGseaTable(RIBOSOMEpaths_Sig_GF.only, ranksGF.only,
              fgsea_C2C5H_GF.only.o, gseaParam = 0.5)
#ggsave(RIBOSOME_GFGSEAplot.only,file=file.path(FigDIR_GF,"RIBOSOME_GFGSEAplot_maxsize200_Clean.ONLLY.pdf"),height=10,width=15)

RIBOpaths_Sig_GF.only <- combpathways[subset(fgsea_C2C5H_GF.only.o[grep("RIBO",fgsea_C2C5H_GF.only.o$pathway),], padj<0.05)$pathway]
RIBO_GFGSEAplot.only <- plotGseaTable(RIBOpaths_Sig_GF.only, ranksGF.only,
              fgsea_C2C5H_GF.only.o, gseaParam = 0.5)
#ggsave(RIBO_GFGSEAplot.only,file=file.path(FigDIR_GF,"RIBO_GFGSEAplot_maxsize200.ONLY.pdf"),height=10,width=15)

# Apoptosis
APOP_NotSig_GF.only <- combpathways[subset(fgsea_C2C5H_GF.only.o[grep("APOP",fgsea_C2C5H_GF.only.o$pathway),], padj<3)$pathway] 
APOP_NotSig_GF_GSEAplot.only <- plotGseaTable(APOP_NotSig_GF.only, ranksGF.only,
              fgsea_C2C5H_GF.only.o, gseaParam = 0.5)
#ggsave(APOP_NotSig_GF_GSEAplot.only,file=file.path(FigDIR_GF,"APOP_GFGSEAplot_maxsize200_notSig.ONLY.pdf"),height=10,width=15) 

## UPR
UNFOLDED_NOTSig_GF.only <- combpathways[subset(fgsea_C2C5H_GF.only.o[grep("UNFOLDED",fgsea_C2C5H_GF.only.o$pathway),], padj<3)$pathway] ## no results
UNFOLDED_NOTSig_GF_GSEAplot.only <- plotGseaTable(UNFOLDED_NOTSig_GF.only, ranksGF.only,
              fgsea_C2C5H_GF.only.o, gseaParam = 0.5)
#ggsave(UNFOLDED_NOTSig_GF_GSEAplot.only,file=file.path(FigDIR_GF,"UNFOLDED_NOTSig_GF_GSEAplot_maxsize200_Padj0.05.ONLY.pdf"),height=10,width=15) 

# Astro
ASTRO_Sig_GF.only <- combpathways[subset(fgsea_C2C5H_GF.only.o[grep("ASTRO",fgsea_C2C5H_GF.only.o$pathway),], padj<0.05)$pathway] 
ASTRO_Sig_GF_GSEAplot.only <- plotGseaTable(ASTRO_Sig_GF.only, ranksGF.only,
              fgsea_C2C5H_GF.only.o, gseaParam = 0.5)
#ggsave(ASTRO_Sig_GF_GSEAplot.only,file=file.path(FigDIR_GF,"ASTRO_Sig_GF_GSEAplot_maxsize200_Padj0.05_Clean_test8501.ONLY.pdf"),height=10,width=15)

# rRNA
rRNA_Sig_GF.only <- combpathways[subset(fgsea_C2C5H_GF.only.o[grep("RRNA",fgsea_C2C5H_GF.only.o$pathway),], padj<0.05)$pathway] 
rRNA_Sig_GF_GSEAplot.only <- plotGseaTable(rRNA_Sig_GF.only, ranksGF.only,
              fgsea_C2C5H_GF.only.o, gseaParam = 0.5)
#ggsave(rRNA_Sig_GF_GSEAplot.only,file=file.path(FigDIR_GF,"rRNA_Sig_GF_GSEAplot_maxsize200_Padj0.05_Clean.ONLY.pdf"),height=10,width=15)


```

## GSEA introduce jitter
```{r}
#ranksGF<- order from highest to lowest
ranksGF.only.s <- sort(ranksGF.only, decreasing = TRUE)

ranksGF.jit <- ranksGF.only.s + runif(length(ranksGF.only.s), min = 1e-10, max = 1e-8)
```

### GSEA to plot multiple plots with correct space, ticks in colour and x axis 
```{r}
plotGseaTable_colored_mirror_with_stats <- function(pathways, stats, fgseaRes,
                                                    col_pos = "#F8766D",
                                                    col_neg = "#00BFC4",
                                                    tick_width = 0.3,
                                                    panel_height = 6,
                                                    spacer_height = 0.3) {

  # Keep only pathways present in the fgsea results
  fgseaRes <- fgseaRes[fgseaRes$pathway %in% names(pathways), ]
  fgseaRes <- fgseaRes[order(fgseaRes$NES, decreasing = TRUE), ]

  plots <- lapply(seq_len(nrow(fgseaRes)), function(i) {
    pw <- fgseaRes$pathway[i]
    genes <- pathways[[pw]]
    gene_hits <- which(names(stats) %in% genes)
    stat_vals <- stats[gene_hits]
    tick_len <- abs(stat_vals) / max(abs(stats))

    pos_idx <- which(stat_vals > 0)
    df_pos <- if (length(pos_idx) > 0) {
      data.frame(x = gene_hits[pos_idx], y = 0, yend = tick_len[pos_idx], col = col_pos)
    } else NULL

    neg_idx <- which(stat_vals < 0)
    df_neg <- if (length(neg_idx) > 0) {
      data.frame(x = gene_hits[neg_idx], y = 0, yend = -tick_len[neg_idx], col = col_neg)
    } else NULL

    hit_df <- rbind(df_pos, df_neg)
    
    n_ranks <- length(stats)
    breaks <- pretty(1:n_ranks, n = 5)
    labels <- format(breaks, big.mark = ",")

    tick_plot <- ggplot(hit_df) +
      geom_segment(aes(x = x, xend = x, y = y, yend = yend, color = col),
                   size = tick_width) +
      scale_color_identity() +
      scale_x_continuous(
    limits = c(1, length(stats)),     # <- COMMON X-AXIS SCALE
    expand = c(0,0),
    breaks = breaks,
    labels = labels) +
      scale_y_continuous(limits = c(-1,1), expand = c(0,0)) +
      theme_classic(base_size = 11) +
      theme(
        axis.title.y = element_blank(),
        axis.title.x = element_text(size = 8, face = "bold", margin = margin(t = 2)),
        axis.text.x = element_text(size = 8, color = "black", margin = margin(t = 0)),
        axis.line.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        plot.margin = margin(0,2,2,2),
        plot.background = element_blank(),
        plot.title = element_text(size = 10, hjust = 0.5)
      ) +
      labs(x = "Rank") +
      ggtitle(pw)

    stat_table <- tableGrob(
      data.frame(NES = round(fgseaRes$NES[i],2), padj = signif(fgseaRes$padj[i],3)),
      rows = NULL,
      theme = ttheme_default(
        core = list(fg_params = list(cex = 0.8), bg_params = list(fill = "white", col = NA)),
        colhead = list(fg_params = list(cex = 0.8), bg_params = list(fill = "white", col = NA))
      )
    )

    combined <- arrangeGrob(tick_plot, stat_table, ncol = 2, widths = c(4,1))
    combined
  })

  # Interleave spacers between pathway plots
  interleaved <- list()
  for(i in seq_along(plots)) {
    interleaved <- c(interleaved, list(plots[[i]]))
    if(i < length(plots)) {
      interleaved <- c(interleaved, list(rectGrob(gp=gpar(col=NA, fill=NA), height=unit(spacer_height,"in"))))
    }
  }

  grid.arrange(grobs = interleaved, ncol = 1, heights = rep(panel_height, length(interleaved)))
}

```

### Function to save multiple GSEA plots in one pdf + list of intersting paths 
```{r}

int.RIBOpaths_Sig_GF.only<-list(
GOCC_PRERIBOSOME= RIBOpaths_Sig_GF.only$GOCC_PRERIBOSOME,
GOBP_RIBOSOMAL_SMALL_SUBUNIT_BIOGENESIS=RIBOpaths_Sig_GF.only$GOBP_RIBOSOMAL_SMALL_SUBUNIT_BIOGENESIS,
GOCC_90S_PRERIBOSOME=RIBOpaths_Sig_GF.only$GOCC_90S_PRERIBOSOME,
GOBP_RIBOSOMAL_LARGE_SUBUNIT_BIOGENESIS=RIBOpaths_Sig_GF.only$GOBP_RIBOSOMAL_LARGE_SUBUNIT_BIOGENESIS,
GOCC_PRERIBOSOME_LARGE_SUBUNIT_PRECURSOR=RIBOpaths_Sig_GF.only$GOCC_PRERIBOSOME_LARGE_SUBUNIT_PRECURSOR)

int.rRNApaths_Sig_GF.only<-list(
REACTOME_RRNA_MODIFICATION_IN_THE_NUCLEUS_AND_CYTOSOL=rRNA_Sig_GF.only$REACTOME_RRNA_MODIFICATION_IN_THE_NUCLEUS_AND_CYTOSOL, 
GOBP_MATURATION_OF_SSU_RRNA= rRNA_Sig_GF.only$GOBP_MATURATION_OF_SSU_RRNA, GOBP_MATURATION_OF_5_8S_RRNA=rRNA_Sig_GF.only$GOBP_MATURATION_OF_5_8S_RRNA, GOBP_RRNA_MODIFICATION=rRNA_Sig_GF.only$GOBP_RRNA_MODIFICATION,
REACTOME_RRNA_PROCESSING=rRNA_Sig_GF.only$REACTOME_RRNA_PROCESSING)

int.Cellcyclepaths_Sig_GF.only<-list(
FISCHER_G1_S_CELL_CYCLE=Cellcyclepaths_Sig_GF.only$FISCHER_G1_S_CELL_CYCLE, WHITFIELD_CELL_CYCLE_G1_S=Cellcyclepaths_Sig_GF.only$WHITFIELD_CELL_CYCLE_G1_S,
GOBP_CELL_CYCLE_CHECKPOINT_SIGNALING=Cellcyclepaths_Sig_GF.only$GOBP_CELL_CYCLE_CHECKPOINT_SIGNALING,
GOBP_CELL_CYCLE_DNA_REPLICATION=Cellcyclepaths_Sig_GF.only$GOBP_CELL_CYCLE_DNA_REPLICATION)

int.int.ASTRO_Sig_GF<-list(LEIN_ASTROCYTE_MARKERS=ASTRO_Sig_GF.only$LEIN_ASTROCYTE_MARKERS, GOBP_ASTROCYTE_DEVELOPMENT=ASTRO_Sig_GF.only$GOBP_ASTROCYTE_DEVELOPMENT)

int.int.APO_NOTSig_GF<-list(REACTOME_APOPTOSIS=APOP_NotSig_GF.only$REACTOME_APOPTOSIS,KEGG_APOPTOSIS=APOP_NotSig_GF$KEGG_APOPTOSIS)


int.int.UPR_NOTSig_GF<-list(GOBP_CELLULAR_RESPONSE_TO_UNFOLDED_PROTEIN=UNFOLDED_NOTSig_GF.only$GOBP_CELLULAR_RESPONSE_TO_UNFOLDED_PROTEIN)


pathway_sets <- list(
    Ribosome = int.RIBOpaths_Sig_GF,
    rRNA = int.rRNApaths_Sig_GF,
    CellCycle = int.Cellcyclepaths_Sig_GF,
    Apoptosis = int.int.APO_NOTSig_GF,
    Unfolded = int.int.UPR_NOTSig_GF,
    Astrocyte = int.int.ASTRO_Sig_GF
    )

plot_multiple_GSEAsets <- function(pathway_list_named,
                                   stats,
                                   fgseaRes,
                                   output_file,
                                   tick_width = 0.6,
                                   panel_height = 6,
                                   width = 6,
                                   height_per_plot = 6
                                   ) {
  
  all_plots <- list()
  
  for (set_name in names(pathway_list_named)) {
    
    p <- plotGseaTable_colored_mirror_with_stats(
      pathways = pathway_list_named[[set_name]],
      stats = stats,
      fgseaRes = fgseaRes,
      tick_width = tick_width,
      panel_height = panel_height
    )
    
    # enforce equal height
    p_fixed <- arrangeGrob(p, heights = unit(1, "null"))
    
    # consistent title for each set
    title_grob <- textGrob(
      paste("Pathway Set:", set_name),
      gp = gpar(fontsize = 16, fontface = "bold")
    )
    
    # Combine title and fixed-size plot into one unit
    combined <- arrangeGrob(title_grob, p_fixed, ncol = 1, heights = c(0.1, 0.9))
    
    all_plots <- c(all_plots, list(combined))
  }
  
  # Save all plots into one PDF, with same panel size
  pdf(file = file.path(FigDIR_GF, "GSEA_NotSig_GF_Comb.Tickplot_corr_space_seed8501_UF.GF.only.pdf"),
      width = width,
      height = height_per_plot * length(all_plots))
  
  grid.arrange(grobs = all_plots,
               ncol = 1,
               heights = rep(unit(1, "null"), length(all_plots)))  # <- uniform height
  
  dev.off()
  
}



plot_multiple_GSEAsets(
  pathway_list_named = pathway_sets,
  stats = ranksGF.jit,
  fgseaRes = fgsea_C2C5H_GF,
  output_file = file.path(FigDIR_TP3, "GSEA_Sig_and_NotSig_GF_Comb.Tickplot_corr_space_Seed8501_UF.GF.ONLY.pdf"),
  tick_width = 0.6,
  panel_height = 6,
  width = 7,
  height_per_plot = 30
)

```
